Performance of Individual Audit Filters in Predicting Opportunities for Improvement

### 2024-10-07 manuscript.Rmd
update manuscript

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
library(webshot2)
```

@@

We conducted a single-centre registry-based retrospective cohort study based on register data collected between 2013 and 2022 from the Karolinska University Hospital in Sweden. Performance was measured using sensitivity, specificity and Cohen’s kappa coefficient for ten audit filters. Bootstrapping with 1,000 iterations calculated 95% confidence intervals.

@@

Our analysis included 8,309 patients. Audit filters showed a Cohen’s kappa coefficient ranging from -0.07 (95% CI [-0.14-0]) for “GCS < 9 and not intubated” and 0.08 (95% CI [0.05-0.12]) for “Massive transfusion”. The Cohen’s kappa coefficient for all audit filters combined was 0.04 (95% CI[0.03-0.04]).

@@

No individual audit filter performed well in predicting opportunities for improvement, which suggests reassessing the current method of screening for mortality and morbidity conferences.

@@

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global] and ranks highest in disability-adjusted life-years
(DALYs) between ages 10-49 years [@vos2020global], impacting
individuals, communities and society at large [@newnam2014impacts; @david2022measuring]. Road
traffic injuries alone cost approximately 2% of gross domestic product
(GDP) in high-income countries [@world2014injuries]. Despite
advancements in trauma care, preventable deaths [@kwon2014pooled;
@ghorbani2018analysis; @teixeira2007preventable; @vioque2014classifying] and errors persist [@nikouline2021errors;
@vioque2014classifying; @o2013opportunities; @teixeira2007preventable]. To address this, trauma quality
improvement programs [@American2022resources], commonly including
mortality and morbidity conferences [@world2009guidelines] have been
implemented. The conferences have shown to foster error reduction and
encourage reporting [@kong2016analysis] without negative feedback
[@lazzara2022morbidity].

Audit filters are criteria used to identify cases for mortality
and morbidity conferences aiming to pinpoint potential opportunities for
improvement [@American2022resources]. Audit filters can involve outcomes
such as "death" or adherence to guidelines, such as "delay to treatment"
[@world2009guidelines]. Opportunities for improvement are typically
associated with failures in initial care [@vioque2014classifying] like
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There seems to be a lack of common, evidence-based
set of audit filters for evaluating trauma care quality
[@stelfox2010quality], and while early studies reported mortality
reduction [@chadbunchachai2001comparative; @chadbunchachai2003study],
newer studies differ [@evans2009audit; @willis2008interpreting]. Moreover, the use of audit filters has been
associated with high frequencies of false positives, ranging from 24% to
80% in different contexts [@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis]. Few
studies have focused on
non-mortality outcome measures, for evaluating performance of audit
filters but the evidence remains inconsistent [@di2008acscot;
@glance2012association] and
some audit filters do not appear to correlate with opportunities for
improvement at all [@lewis2017delay]. However, the perceived
usefulness of audit filters remains high [@berg2022perceived]. The aim of this study is to determine the performance of individual
audit filters in predicting
opportunities for improvement and their relevance as a
screening tool for mortality and morbidity conferences

@@

The trauma centre at the Karolinska University Hospital also keeps a
local care quality database with information intended for the mortality
and morbidity conferences, including identified triggered audit filters
and opportunities for improvement. Opportunities for improvement are
assessed in multiple stages of scrutiny. Notably, mortality is directly
referred to the mortality and morbidity conferences to assess
opportunities for improvement.

@@

The study included all patients treated at Karolinska University
Hospital from 2013 to 2022 who underwent screening for opportunities for
improvement. Patients aged under 15 years were excluded due to differences in
clinical management.

@@

The ten audit filters used at the Karolinska University Hospital served
as the exposure variables. For each audit filter, we identified the cohort of patients in which the filter was applicable, and evaluated the filters’ performance in that cohort: For “ISS > 15 and no team activation” and “ISS > 15 and not in ICU” we evaluated the filters’ performance in patients with an ISS greater than 15; for "GCS < 9 and not intubated" we analyzed patients with a GCS less than 9; and lastly, for "No anticoagulants within 72 hours after TBI" we identified patients with head trauma as defined by the precence of at least one head injury according to available AIS codes.

The data was retrieved from the Karolinska Trauma Registry and trauma
care quality database. Patient data on vital signs, care processes and
interventions, level of care and time aspects was retrieved from the trauma registry whereas the exposure, audit filters and the outcome, and opportunities
for improvement were retrieved from the quality database.

@@

We used R [@R2022language] for statistical analysis. Patient
characteristics for cases with and without opportunities for improvement
were compared using the Wilcoxon signed-rank test for continuous
variables and the Chi-squared test for categorical variables. The
predictive performance of audit filters was assessed by calculating the
sensitivity, specificity and Cohen's kappa for each audit filter in
predicting opportunities for improvement. Bootstrapping with 1,000
iterations was applied to estimate 95% confidence intervals for all
performance metrics. We selected cohorts relevant to each filter: For
"ISS \> 15 and no team activation" and "ISS \> 15 and not in ICU" we
examined patients with an ISS greater than 15. For "GCS \< 9 and not
intubated," we focused on patients with a GCS less than 9. Lastly, for
"No anticoagulants within 72 hours after TBI" we identified patients
with head trauma as defined by the presence of at least one head injury
according to available AIS codes. Missing values were numbered but
ignored in the analysis.

@@

Patients with an opportunity for improvement had significantly higher
median age,
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2)` years vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1)` years,
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")`, median 
ISS `r inline_text(tableOne, variable = ISS, column = stat_2)` vs
`r inline_text(tableOne, variable = ISS, column = stat_1)`,
`r inline_text(tableOne, variable = ISS, column = "p.value")` and
intubations in the emergency department (ED), 
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)` patients
vs
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1)` patients,
`r inline_text(tableOne, variable = ed_intubated, column = "p.value")`.
Patients with opportunity for improvement had longer times to first major intervention from hospital arrival median
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2)`
minutes vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1)` minutes,
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`.
and longer times to first CT, median
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2)` minutes vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1)` minutes,
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")` minutes.
The frequency of opportunities for improvement was highest in patients
admitted to the intensive care unit,
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
patients, compared to general ward
`r inline_text(tableOne, variable = host_care_level, level = "General ward", column = stat_2)`. Notably, there was a statistically significant
correlation between GCS at ED and opportunities for improvement (p \<
0.001).
Gender, mean systolic blood pressure, death at 30 days and trauma team
activation were not significantly correlated with opportunities for
improvement (p\>0.05). The highest amount of missing data was observed for the audit
filter "time to first major intervention", which was missing for 5,990 patients, (72.1%). See Table 2
for details of selected patient characteristics.

@@

The audit filter with the highest sensitivity was "\> 60 min until first
intervention" (83.3%, 95% CI [79.2-87.5]). The lowest sensitivity was
seen in "SBP \< 90" (8%, 95% CI [5.6-10.2]) and "Dead at 30 days" (8%,
95% CI [5.6-10.2]). The audit filter with the highest specificity was
"Massive transfusion" (95.8%, 95% CI [95.4-96.3]). The lowest
specificity was seen in "ISS \> 15 and no team activation" (21.6%, 95%
CI [19.9-23.2]). The sensitivity for all audit filters combined was
92.6% (95% CI [90.5-95]) and the specificity was 29.4% (95% CI
[28.4-30.5]).

The highest Cohen's kappa was seen for the audit filter "Massive
transfusion" (0.08, 95% CI [0.05-0.12]). The smallest Cohen's kappa was
seen in "GCS \< 9 and not intubated" (-0.07, 95% CI [-0.14-0]. The
Cohen´s kappa for all audit filters combined was 0.04 (95% CI
[0.03-0.04]. See table 3 for the performance of all audit filters using
sensitivity, specificity and Cohen's kappa coefficient.

@@

agree

disagree
Some previous studies have shown similar findings to this study. Lewis et al. looked at one audit filter, "delay to operating room" of more than 2 hours and its correlation with opportunities for improvement. They found that the audit filter had no significant ability to predict opportunities for improvement. Similarly, the delay-related audit filters: "\>60 min until first intervention" and
"\>30min until first CT" showed noability to predict opportunities for improvement. A big meta analysis by Evans et al. sought to identify a common set of evidence based audit filters but no study fulfilled the inclusion criteria.
A 2008 study by Di Bartolomeo et al. [@di2008acscot] looked at 4 ASCOT audit filters in relation to 30-day mortality and a morbidity measure. One audit filter showed a correlation with mortality but no audit filter showed a correlation with the morbidity measure.

A recent study by Berg et al. [@berg2024effects], looking at implementation of audit filters on mortality concluded that audit filters might be useful in reducing the mortality. 
Previous studies suggest poor performance of audit filters for improving
trauma care [@evans2009audit; @stelfox2010quality; @willis2008interpreting; @di2008acscot; @glance2012association; @lewis2017delay]. This study adds to that evidence by showing that audit filters perform poorly in predicting opportunities for
improvement encompassing mortality and non-mortality
associated outcomes. Some audit filters might be more reflective of clinical decisions rather than quality of care [@glance2012association]. For instance, an abnormal GCS because of intoxication
is managed differently to an abnormal GCS due to head injury, which could explain the poor performance of "GCS < 9 and not intubated", similarly shown by
Willis et al. [@willis2008interpreting]. This could also explain the poor performance of the delay-related audit filters: "\>60 min until first intervention" and
"\>30min until first CT". Similarly, Lewis et al. [@lewis2017delay] showed that delay
to the operating room was not associated with opportunities for improvement.
In contrast, Teixeira et al. [@teixeira2007preventable] identified delays to
treatment and delays to CT as common errors in trauma care. A recent
study in Japan by Yamamoto et. al [@yamamoto2023immediate] observed
significantly reduced mortality with whole-body CT scans within 10
minutes of arrival to the ED. If anything, this highlights the unique opportunities for improvement in different hospitals. The audit filters also varied greatly in terms of expected mortality. Some audit
filters such as "Massive transfusion" and "Dead at 30 days" have a high
expected mortality. In such cases the reduced time frame for any potential
opportunity for improvement to occur could explain their poor performance. When audit filters were combined they had a low specificity relative to its sensitivity which corresponds to previous studies showing high numbers of false positives [@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis]. The sensitivity and specificity for individual audit filters varied substantially, suggesting that some audit filters are too rigid, missing the majority of opportunities for improvement while some are too inclusive, causing high numbers of false positives [@willis2008interpreting].

There were some limitations in this study. Firstly, it is unlikely that the specific associations presented in this study are generalizable to other hospitals with similar systems, but the broader finding that audit filters are poorly associated with opportunities for improvement is likely to be externally valid [@world2009guidelines]. Secondly, instances where the
same patient case had multiple audit filters introduce uncertainty
regarding which specific audit filter truthfully predicted an
opportunity for improvement. Finally, the number of missing values
differed significantly among the audit filters, resulting in varying
degrees of validity for each filter.

The static nature of audit filters inherently limit their adaptability to
emerging healthcare challenges and resolved quality gaps, reducing their
utility over time. Therefore, regular evidence-based reviews are needed
to ensure performance [@santana2014development], though such reviews may
be cost-prohibitive. Furthermore, they might overshadow other
contributing factors to opportunities for improvement
[@willis2008interpreting].

In summary, audit filters are widely used as a tool in quality
improvement programmes and should reflect current gaps in trauma care
quality. In this study, no individual audit filter or their combination were
successful in predicting opportunities for improvement. Given these
results, future studies should focus on developing new tools with higher predictive
ability.

>>>>

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
library(webshot2)

```

@@

We conducted a single-centre registry-based retrospective cohort study based on register data collected between 2013 and 2022 from the Karolinska University Hospital in Sweden. Performance was measured using sensitivity, specificity and Cohen’s kappa coefficient for ten audit filters.

@@

Our analysis included 8,309 patients. Audit filters demonstrated inconsistent agreement between filters and opportunities for improvement with a Cohen’s kappa coefficient ranging from -0.07 (95% confidence interval [CI], -0.14-0) for “GCS < 9 and not intubated” and 0.08 (CI, 0.05-0.12) for “Massive transfusion”. The Cohen’s kappa coefficient for all audit filters combined was 0.04 (CI, 0.03-0.04).

@@

No individual audit filter performed well in predicting opportunities for improvement, which suggests that the current method of screening for mortality and morbidity conferences should be revisited.

Level III, Prognostic/Epidemiological
ofi, audit filter, quality improvement

@@

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global] and ranks highest in disability-adjusted life-years
(DALYs) between ages 10-49 years [@vos2020global], impacting
individuals, communities and society at large [@newnam2014impacts; @david2022measuring]. Road
traffic injuries alone cost approximately 2% of gross domestic product
(GDP) in high-income countries [@world2014injuries]. Despite
advancements in trauma care, preventable deaths [@kwon2014pooled;
@ghorbani2018analysis; @teixeira2007preventable; @vioque2014classifying] and avoidable medical errors persist [@nikouline2021errors;
@vioque2014classifying; @o2013opportunities; @teixeira2007preventable]. To address this, trauma quality
improvement programs [@American2022resources], commonly including
mortality and morbidity conferences [@world2009guidelines] have been
implemented. The conferences have shown to foster error reduction and
encourage reporting [@kong2016analysis] without negative feedback
[@lazzara2022morbidity].

Audit filters are criteria used to identify cases for mortality
and morbidity conferences aiming to pinpoint potential opportunities for
improvement [@American2022resources]. Audit filters can involve outcomes
such as "death" or adherence to guidelines, such as "delay to treatment"
[@world2009guidelines]. Opportunities for improvement are typically
associated with failures in initial care [@vioque2014classifying] like
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There is a lack of common, evidence-based
set of audit filters for evaluating trauma care quality
[@stelfox2010quality], and while early studies reported mortality
reduction [@chadbunchachai2001comparative; @chadbunchachai2003study],
newer studies differ [@evans2009audit; @willis2008interpreting]. Moreover, the use of audit filters has been
associated with high frequencies of false positives, ranging from 24% to
80% in different contexts [@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis]. Few
studies have focused on
non-mortality outcome measures, for evaluating performance of audit
filters but the evidence remains inconsistent [@di2008acscot;
@glance2012association] and
some audit filters do not appear to correlate with opportunities for
improvement at all [@lewis2017delay]. However, the perceived
usefulness of audit filters remains high [@berg2022perceived]. The aim of this study is to determine the performance of individual
audit filters in predicting
opportunities for improvement and their relevance as a
screening tool for mortality and morbidity conferences

@@

The trauma centre at the Karolinska University Hospital also keeps a
local care quality database with information intended for the mortality
and morbidity conferences, including identified triggered audit filters
and opportunities for improvement. Opportunities for improvement are
assessed in multiple stages of scrutiny. Notably, patient deaths are always reviewed to determine if they could have been prevented and to identify
opportunities for improvement.

@@

The study included all patients treated at Karolinska University
Hospital from 2013 to 2022 and who were screened for opportunities for
improvement. Patients aged under 15 years were excluded.

@@

The ten audit filters used at the Karolinska University Hospital served
as the exposure variables. For each audit filter, we identified the cohort of patients in which the filter was applicable, and evaluated their performance in that cohort: For “ISS > 15 and no team activation” and “ISS > 15 and not in ICU” we evaluated the filters’ performance in patients with an ISS greater than 15; for "GCS < 9 and not intubated" we analysed patients with a GCS less than 9; for "> 30 min until first intervention" and "> 60 min until first CT" we selected patients who underwent a first intervention and those who received a CT respectively. Lastly, for "No anticoagulants within 72 hours after TBI" we identified patients with head trauma as defined by the presence of at least one head injury according to available AIS codes.

The data was retrieved from the Karolinska Trauma Registry and trauma
care quality database. Patient data on vital signs, care processes and
interventions, level of care and time aspects was retrieved from the trauma registry whereas the exposure, audit filters and the outcome, and opportunities
for improvement were retrieved from the trauma care quality database.

@@

We used R [@R2022language] for statistical analysis. Patient
characteristics for cases with and without opportunities for improvement
were compared using the Wilcoxon signed-rank test for continuous
variables and the Chi-squared test for categorical variables. The
predictive performance of audit filters was assessed by calculating the
sensitivity, specificity and Cohen's kappa for each audit filter in
predicting opportunities for improvement. A Cohen's kappa of 0-0.2 was interpreted as no reliability for the predictions while a value above 0.2 was interpreted as having a minimal reliability up to perfect reliability being 1 [@mchugh2012interrater]. Bootstrapping with 1,000
iterations was applied to estimate 95% confidence intervals for all
performance metrics. We selected cohorts relevant to each filter: For
"ISS \> 15 and no team activation" and "ISS \> 15 and not in ICU" we
examined patients with an ISS greater than 15. For "GCS \< 9 and not
intubated," we focused on patients with a GCS less than 9. Lastly, for
"No anticoagulants within 72 hours after TBI" we identified patients
with head trauma as defined by the presence of at least one head injury
according to available AIS codes. Missing values were numbered but
ignored in the analysis for example those without a registered blood pressure was ignored when analysing "SBP < 90".

@@

Patients with an opportunity for improvement had significantly higher
median age,
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")` years vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")` 
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")`) years, median 
ISS `r inline_text(tableOne, variable = ISS, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")` vs
`r inline_text(tableOne, variable = ISS, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = ISS, column = "p.value")`) and
intubations in the emergency department (ED), 
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)` patients
vs
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1, pattern = "{N} ({p}%;")`
`r inline_text(tableOne, variable = ed_intubated, column = "p.value")`) patients.
Patients with opportunity for improvement had longer times to first major intervention from hospital arrival median
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")`
minutes vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")` 
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`) minutes.
and longer times to first CT, median
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")` minutes vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")`) minutes.

The frequency of opportunities for improvement was highest in patients
admitted to the intensive care unit,
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
patients, compared to general ward
`r inline_text(tableOne, variable = host_care_level, level = "General ward", column = stat_2)`. Notably, there was a statistically significant
correlation between GCS at ED and opportunities for improvement (p \<
0.001).
Gender, mean systolic blood pressure, death at 30 days and trauma team
activation were not significantly correlated with opportunities for
improvement (p\>0.05). The highest amount of missing data was observed with "time to first major intervention", which was missing for 5,990 patients, (72.1%), partly because the absence of an initial CT or major intervention due to the clinical management being observation resulted in their registration as missing. See Table 2
for details of selected patient characteristics.

@@

The audit filter with the highest sensitivity was "\> 60 min until first
intervention" (83.3%, 95% CI, 79.2-87.5). The lowest sensitivity was
seen in "SBP \< 90" (8%, 95% CI, 5.6-10.2) and "Dead at 30 days" (8%,
95% CI, 5.6-10.2). The audit filter with the highest specificity was
"Massive transfusion" (95.8%, 95% CI, 95.4-96.3). The lowest
specificity was seen in "ISS \> 15 and no team activation" (21.6%, 95%
CI, 19.9-23.2). The sensitivity for all audit filters combined was
92.6% (95% CI, 90.5-95) and the specificity was 29.4% (95% CI, 28.4-30.5).

The highest Cohen's kappa was seen for the audit filter "Massive
transfusion" (0.08, 95% CI, 0.05-0.12). The smallest Cohen's kappa was
seen in "GCS \< 9 and not intubated" (-0.07, 95% CI, -0.14-0. The
Cohen´s kappa for all audit filters combined was 0.04 (95% CI, 0.03-0.04). See table 3 for the performance of all audit filters using
sensitivity, specificity and Cohen's kappa coefficient.

@@

Few studies have examined the link between individual audit filters and opportunities for improvement. Given the myriad of different audit filters it is challenging to accurately compare our findings with studies using a completely different set of audit filters. The delay-related audit filters in this study: "\>60 min until first intervention", 
"\>30 min until first CT" did not predict opportunities for improvement. Interestingly, time to definitive treatment and time to first CT was associated with opportunities for improvement in the patient characteristics analysis similar to a previous study by Albaaj et al. [@albaaj2023patient] on patient factors associated with opportunities for improvement. However, while statistically significant, the observed effect size suggests low clinical relevance. Similarly, one study by Lewis et al. [@lewis2017delay] found that delays to operating room exceeding 2 hours from emergency department arrival did not individually predict opportunities for improvement.

Furthermore, the audit filter "GCS < 9 and not intubated" showed no correlation with opportunities for improvement in our study. Comparably, Willis et al. [@willis2008interpreting] found no association between "GCS below 9 and not intubated" and mortality or length of hospital stay but it was associated with shorter stays in the ICU. Contrary, Glance et. al [@glance2012association] found that "No definitive airway in patient with GCS score ≤ 8" was associated with increased mortality risk but reduced risk of major complication and death. This was concluded as reflective of clinical decision rather than quality of care [@willis2008interpreting; @glance2012association]. For instance, an abnormal GCS because of intoxication
is managed differently to an abnormal GCS due to head injury. Furthermore, Willis et al. found that "failure to activate trauma team at major trauma service" was not associated with poor outcomes, similar to our findings.

Di Bartolomeo et al. [@di2008acscot] looked at the association between four audit filters ("Ambulance scene time >20 min", "Patient with a Glasgow coma
scale score (GCS) < 13 who does not receive a CT scan of the head within 2 h of arrival at the hospital", "Patients with epidural or subdural haematoma receiving craniotomy >4 h after arrival at ED, excluding those performed for intracranial pressure monitoring", "Abdominal,
thoracic, vascular or cranial surgery performed >24 h after arrival") and 30-day mortality and 6-month disability. One audit filter: "abdominal, thoracic, vascular or cranial surgery
performed >24 h after arrival" showed contradictory outcomes with reduced mortality but increased disability. This may be explained by aggressive trauma care producing additional survivors with poor functional outcomes or underestimating the mortality rate beyond 30 days. The rest of the audit filters showed no significant association with outcomes.

A study by Chadbunchachai et al. [@chadbunchachai2001comparative] reported a reduction in preventable deaths following the implementation of audit filters, with similar results in a subsequent study [@chadbunchachai2003study]. In 2009 Evans et al. [@evans2009audit] conducted a review to evaluate the effectiveness of audit filters in improving trauma care quality but found no studies of sufficient methodological quality for inclusion. The studies by Chadbunchachai et al. were excluded due to insufficient pre- and post-implementation data points. They concluded that there is no high-quality evidence supporting the use of audit filters in improving trauma care. Similarly, Stelfox et al. [@stelfox2010quality] found no clearly defined, evidence-based audit filters. A more recent study by Berg et al. [@berg2024effects] also observed decreased mortality after the implementation of audit filters in India. Furthermore, the perceived usefulness of audit filters in trauma quality improvement among health care professionals remains high [@berg2022perceived; @santana2014development].

There are some limitations in this study. Firstly, it is unlikely that the specific associations presented in this study are generalizable to other hospitals with similar systems [@di2008acscot; @glance2012association], but the broader finding that audit filters are poorly associated with opportunities for improvement is likely to be externally valid [@world2009guidelines]. Secondly, instances where the
same patient case had multiple audit filters flagged introduce uncertainty
regarding which specific audit filter truthfully predicted an
opportunity for improvement, introducing a level of confounding. Finally, the number of missing values
differed significantly among the audit filters, resulting in varying
degrees of validity for each audit filter.

The static nature of audit filters inherently limit their adaptability to
emerging healthcare challenges and resolved quality gaps, reducing their
utility over time [@glance2012association]. Therefore, regular evidence-based reviews are needed
to ensure performance [@santana2014development], though such reviews may
be cost-prohibitive. Furthermore, they may overshadow other
contributing factors to opportunities for improvement, i.e confounding variables [@glance2012association; @willis2008interpreting].

In summary, audit filters are widely used as a tool in quality
improvement programmes and should reflect current gaps in trauma care
quality. In this study, no individual audit filter or their combination were
successful in predicting opportunities for improvement. Given these
results, in combination with the lack of high-quality evidence showing that implementing audit filters improve patient outcomes, future studies should focus on developing new tools with higher predictive
ability to identify opportunities for improvement.
